\chapter{Site requirements}

\section{Sensor unit requirements}
The sensor unit should be located outside away from human
distrubances. The site for the sensor unit should be chosen with
regard to the following requirements, with the highest priority given
first.

\begin{itemize}
\item Within range of the base unit.
\item Away from moving metal objects, for example, trains (more than
  \SI{50}{\metre}), cars (more than \SI{20}{\metre}) and garage doors.
\item Away from static metal objects, in particular those containing
  the \emph{ferro-magnetic} materials iron, nickel and cobalt.
\end{itemize}

The FLC100 fluxgate magnetometer sensor is slightly sensitive to
temperature variations. To ensure correct behaviour a stabilised
temperature environment is required. This is obtained by burying the
sensor.

\section{Base unit requirements}

The base unit (Raspberry Pi and radio module) requires mains power and
a wired network connection. Wireless networking is also possible but
has not been tested; follow the instructions on the Raspberry Pi
website. The base unit should be located indoors and as close to the
sensor unit as possible.

\subsection{Network requirements}

The following network requirements are needed for the system to
operate fully:
\begin{itemize}
\item \dns\ resolution. This is normally provided
  as standard on most networks.
\item Outgoing access on port 80 (\http) and 443 (\https). Required
  for software updates.
\item Outgoing access on port 123 (\ntp), or access to a local \ntp\
  server.
\item Outgoing \ssh\ access (port 22).
\end{itemize}




